---
title: "Introduction to simone"
author: "Chiquet, Smith, Grasseau, Matias, Ambroise"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to simone}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 5, fig.height = 5)
```

## Overview

The **simone** package implements Statistical Inference for MOdular NEtworks (SIMoNe).
It infers gene co-expression networks based on partial correlation coefficients from
microarray experiments, using a Gaussian graphical model (GGM) with adaptive penalization
driven by a latent modular structure.

**Reference**: Chiquet J., Smith A., Grasseau G., Matias C., Ambroise C. (2008).
SIMoNe: Statistical Inference for MOdular NEtworks.
*Bioinformatics*, 25(3), 417–418. doi:[10.1093/bioinformatics/btn637](https://doi.org/10.1093/bioinformatics/btn637)

---

### Simulation of one data set ($n=100$)

```{r, message=FALSE}
library(simone)
set.seed(2)
p <- 11
g <- rNetwork(p, pi = 0.2, name = "Theoretical Graph")
plot(g)
True.correlation.matrix <- cov2cor(solve(g$Theta))
data <- rTranscriptData(n = 100, g, sigma = 0)
```

### Computation of the observed correlation matrix

```{r}
Rho <- cor(data$X)
image(Rho)
# mean square diff between truth and estimation
mean((Rho - True.correlation.matrix)^2)
# How many independences?
sum(Rho == 0)
```

All pairs of variables are dependent since no correlation is set to 0. So let us
compute the partial correlations via the concentration matrix.

### Infer a network with a penalty parameter equal to 0.3

```{r, echo=TRUE, message=FALSE}
infered.net <- simone(data$X, control = setOptions(penalties = 0.3))
confusion_matrix <- function(truth, estimation) {
  up <- upper.tri(truth, diag = FALSE)
  graph_theorique <- factor(as.numeric(!truth == 0), levels = c(0, 1))
  graph_infere    <- factor(as.numeric(!estimation == 0), c(0, 1))
  return(table(infere = graph_infere[up], theorique = graph_theorique[up]))
}
(m <- confusion_matrix(truth = g$A, estimation = infered.net$networks[[1]]))
```

From the confusion matrix we see that the algorithm has 38 TP and 8 TN but does
make 9 mistakes.

### Using different penalties

```{r}
## Running simone
res <- simone(data$X, type = "steady-state")

## Plotting the results
plot(res, ref.graph = g$A, output = "ROC", ask = FALSE)
```

The best network according to BIC is the following. Let us plot it with its
confusion matrix.

```{r}
res <- simone(data$X)
selected.net <- getNetwork(res)
plot(selected.net)
confusion_matrix(g$A, selected.net$A)
```

### The flow-cytometry dataset

The Sachs *et al.* protein signalling dataset is available from the Stanford
ElemStatLearn archive. The chunk below is not evaluated during package build
(requires internet access), but can be run interactively.

```{r sachs, eval=FALSE}
source <- "https://web.stanford.edu/~hastie/ElemStatLearn/datasets/sachs.data"
protein_data <- read.table(url(source),
  col.names = c("praf","pmek","plcg","PIP2","PIP3","p44/42",
                "pakts473","PKA","PKC","P38","pjnk"))

grid_lambda <- c(0, 0.01, 0.05, 0.1, 0.5, 1)
op <- par(mfrow = c(2, 3))
for (i in seq_along(grid_lambda)) {
  infered.net <- simone(protein_data, control = setOptions(penalties = grid_lambda[i]))
  plot(getNetwork(infered.net))
}
par(op)
```

The best network according to the BIC criterion is the following:

```{r sachs-best, eval=FALSE}
res <- simone(protein_data)
plot(getNetwork(res))
```
